Discriminating Seagrasses From Green Macroalgae in European Intertidal areas using high resolution multispectral drone imagery.

### 2024-02-27 _manuscript/_tex/index.tex
add all big files

\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\usepackage{bookmark}

@@

\begin{abstract}
In September 2021, a significant jump in seismic activity on the island
of La Palma (Canary Islands, Spain) signaled the start of a volcanic
crisis that still continues at the time of writing. Earthquake data is
continually collected and published by the Instituto Geográphico
Nacional (IGN). \ldots{}
\end{abstract}

@@

@@

  \bibliography{library.bib}

>>>>

\usepackage{fontspec}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{colortbl}
\usepackage{hhline}
\newlength\Oldarrayrulewidth
\newlength\Oldtabcolsep
\usepackage{longtable}
\usepackage{array}
\usepackage{hyperref}
\usepackage{float}
\usepackage{wrapfig}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\usepackage{bookmark}

@@

\begin{abstract}
Coastal areas host seagrass meadows, which offer crucial ecosystem
services including erosion control and carbon sequestration. However,
these areas are increasingly impacted by human activities, leading to
seagrass decline and habitat fragmentation. In situ surveys,
traditionally performed to monitor these ecosystems face limitations on
temporal and spatial coverage, particularly in intertidal zones,
prompting the use of satellite data within monitoring programs. Yet,
satellite remote sensing struggles with spatial and spectral resolution,
making it difficult to discriminate seagrass from other macrophytes in
highly heterogenous meadows. To address these challenges, drone images
acquired with multi-spectral sensors offer a promising solution. This
study focuses on using drones acquisitions for mapping intertidal
macrophytes, effectively discriminating between seagrass and green
macroalgae. Ten drone flights were conducted at two different altitudes
(12m and 120m) across diverse European habitats in France and Portugal.
Low altitude flights were used to train a Deep Learning classifier based
on Neural Networks to discrimintate among 5 intertidal vegetation
classes. Drone mapping demonstrated an overall accuracy of 94\% across
all the sites and images, covering a total area of 467 000 m². The model
exhibited an accuracy of 96.4\% in identifying seagrass.
\end{abstract}

@@

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

@@

\section{Material \& Methods}\label{material-methods}

\subsection{Study sites}\label{study-sites}

Seven study sites distributed between France and Portugal were selected
for their relatively extensive intertidal seagrass beds. Two sites are
located in Gulf of Morbihan (Figure~\ref{fig-map} A), France (47.5791°N,
2.8018°W). This gulf covers an area of 115 km² and is only connect to
the sea through a 900m wide channel. A total of 53 small islands are
scattered across the gulf leading to 250 km of shorelines. Patchy
seagrass meadows can be found on a lot of these islands. Within the Gulf
of Morbihan we have chosen two sites : One is located in one of these
island (Arz island). The other one is located in the sourthern part of
the Gulf (Duer). Two others sites are located in Bourgneuf Bay, France
(46.9849°N, 2.1488°W). This bay is a semi-enclosed macrotidal bay,
protected from waves by Noirmoutier Island. With up to 6m in tidal range
during spring tide, a 6 km² seagrass meadow can uncover at low tide. The
sites we surveyed (L'Epine and Barbatre, Figure~\ref{fig-map} B) are
located in the southernmost part of the bay, along the Noirmoutier
Island, and host a monospecific \emph{Zostera noltei} seagrass bed with
very little mixing with other macrophytes. Three sites have been survey
in the Ria de Aveiro Lagoon in Portugal (40.6887°N, 8.6810°W). The
extent of this lagoon is about 80 km² with many narrow channels, large
salt marshes and many mudflats that uncovers at low tide. It is
connected to the open sea through a single channel, with a tidal lag
between the North and the South of the lagoon. The southernmost site
(Gafanha) is a mudflat located in an affluent of the lagoon whereas the
two other sites are situated in the middle of the lagoon and are only
accessible by boat (Figure~\ref{fig-map} C). These Portuguese sites were
characterized by a more diverse intertidal area, where the seagrass
meadow could intermingle with red algae (Rhodophyceae), brown algae
(Phaeophyceae), or green algae (Chlorophyceae).

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\phantomsection\label{cell-fig-map}
\begin{figure}[H]

\centering{

\includegraphics[width=26.67in,height=\textheight]{./Figures/Fig1_Map_Drone_Sites.png}

}

\caption{\label{fig-map}Sites where the different drone flights were
made. A: Gulf of Morbihan, B: Bourngeuf Bay, C: Ria de Aveiro Lagoon.
Light green represent terrestrial areas whereas darkgreen represent
intertidal areas.}

\end{figure}%

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\subsection{Drone Data}\label{drone-data}

At each location, a DJI Matrice 200 quadcopter drone equipped with a
Micasense RedEdge Dual MX multispectral camera was flown to take 1.2
million pixel reflectance photographs in 10 spectral bands, from blue
(444 nm) to near infrared (840 nm). An angle of 90° has been maintained
between the sun and the drone's route to guarantee uniform lighting
conditions between flight lines. A side overlap of 70\% and a front
overlap of 80\% between each image has been set. Images have been
calibrated in reflectance using a calibration panel reflective at
\textasciitilde50\% provided by the camera's manufaturer and a
Downwelling Light Sensor (DLS2) that was utilized to acquire irradiance
data during the flight. A Structure-from-Motion photogrammetry software
(Agisoft Metashape) has been used to process images to obtain
multispectral orthomosaics of each flight. The workflow for
orthomosaicking has remained consistent for every flight. First, tying
key points were detected inside of each image and between overlapping
images in order to obtain a sparse point cloud. This cloud was cleaned
using reprojection accuracy metric in order to remove noisy points. A
dense point cloud has been then produced using a structure from motion
algorithm. A surface interpolation of this dense point cloud has been
made to obtain a Digital Surface Model (DSM), used to reconstruct the
multispectral ortho-image. Across all the sites, flights were made at
two different altitudes : 12 m or/and 120 m. Low altitude drone flights
produce ortho-images with a very high spatial resolution (8 mm per
pixel), making it simple to visually distinguish between the various
types of vegetation. High altitude fights on the other hand allow to
cover large areas and produced images with a pixel size of 80 mm (Table
1).

\global\setlength{\Oldarrayrulewidth}{\arrayrulewidth}

\global\setlength{\Oldtabcolsep}{\tabcolsep}

\setlength{\tabcolsep}{0pt}

\renewcommand*{\arraystretch}{1.5}

\providecommand{\ascline}[3]{\noalign{\global\arrayrulewidth #1}\arrayrulecolor[HTML]{#2}\cline{#3}}

\begin{longtable}[c]{ccccc}
\caption{List of drone Flight, summarising the date, the altitude and the purpose
of each flight.}\tabularnewline

\multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Country}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Site}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 1.5pt}>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Name}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Altitude}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Utility}}} \\

\ascline{1.5pt}{000000}{1-5}\endfirsthead

\multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Country}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Site}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 1.5pt}>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Name}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Altitude}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Utility}}} \\

\ascline{1.5pt}{000000}{1-5}\endhead

\multicolumn{1}{>{}c}{} & \multicolumn{1}{>{}c}{} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 1.5pt}>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Arz\ Island}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{12m}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Training}}} \\

\ascline{1.5pt}{BEBEBE}{3-5}

\multicolumn{1}{>{}c}{} & \multicolumn{1}{>{}c}{} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 1.5pt}>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Duer}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{12m}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Training}}} \\

\ascline{1.5pt}{BEBEBE}{3-5}

\multicolumn{1}{>{}c}{} & \multicolumn{1}{>{}c}{\multirow[c]{-3}{*}{\parbox{0.75in}{\centering \textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Morbihan}}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 1.5pt}>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Duer}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{120m}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Validation}}} \\

\ascline{1.5pt}{666666}{2-5}

\multicolumn{1}{>{}c}{} & \multicolumn{1}{>{}c}{} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 1.5pt}>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Barbâtre}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{120m}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Validation}}} \\

\ascline{1.5pt}{BEBEBE}{3-5}

\multicolumn{1}{>{}c}{\multirow[c]{-5}{*}{\parbox{0.75in}{\centering \textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{France}}}}} & \multicolumn{1}{>{}c}{\multirow[c]{-2}{*}{\parbox{0.75in}{\centering \textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Bourgneuf\ Bay}}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 1.5pt}>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{L'Epine}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{120m}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Validation}}} \\

\ascline{1.5pt}{000000}{1-5}

\multicolumn{1}{>{}c}{} & \multicolumn{1}{>{}c}{} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 1.5pt}>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Inner\ Lagoon}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{120m}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Validation}}} \\

\ascline{1.5pt}{BEBEBE}{3-5}

\multicolumn{1}{>{}c}{} & \multicolumn{1}{>{}c}{} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 1.5pt}>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Mataducos}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{120m}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Validation}}} \\

\ascline{1.5pt}{BEBEBE}{3-5}

\multicolumn{1}{>{}c}{} & \multicolumn{1}{>{}c}{} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 1.5pt}>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Gafanha}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{120m}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Validation}}} \\

\ascline{1.5pt}{BEBEBE}{3-5}

\multicolumn{1}{>{}c}{\multirow[c]{-4}{*}{\parbox{0.75in}{\centering \textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Portugal}}}}} & \multicolumn{1}{>{}c}{\multirow[c]{-4}{*}{\parbox{0.75in}{\centering \textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Aveiro\ Lagoon}}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 1.5pt}>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Gafanha}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{12m}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Training}}} \\

\ascline{1.5pt}{000000}{1-5}

\end{longtable}

\arrayrulecolor[HTML]{000000}

\global\setlength{\arrayrulewidth}{\Oldarrayrulewidth}

\global\setlength{\tabcolsep}{\Oldtabcolsep}

\renewcommand*{\arraystretch}{1}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\subsection{Field sampling}\label{field-sampling}

Before each flight, targets used as Ground Control Points (GCPs) were
dispersed and georeferenced with a Trimble © Geo XH 6000 differential
GPS (dGPS). GCPs were used to correct georeferencing imprecision of
orthomosaics with an horizontal and vertical accuracy of 10cm. dGPS was
also used to georeference quardrats of 0.25cm² used to assess the
presence or absence of 5 key classes of vegetation : Bacillariophyceae
(Diatoms forming large biofilm over the sediment during low tide),
Phaeophyceae (Brown algae), Magnoliopsida (Seagrass), Chlorophyceae
(Green algae) and Rhodophyceae (Red algae)
(Figure~\ref{fig-vegetation}). Picture of each quadrat were taken and
uploaded on the online on the Global Biodiversity Information Facility
(GBIF) platform \citep{BedeGbif}. Each photograph was also processed to
estimate the percent cover of each type of vegetation. Georeferenced
hyperspectral reflectance signatures of each vegetation class were also
sampled using an ASD FieldSpec HandHeld 2 spectroradiomete, which
acquires reflectance between 325 and 1075 nm, with 1 nm of spectral
resolution. The methodological scheme followed in this study is
presented in Figure~\ref{fig-workflow}.

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\phantomsection\label{cell-fig-vegetation}
\begin{figure}[H]

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{Figures/Spectral_shapes_total.png}

}

\caption{\label{fig-vegetation}Class of vegetation used to train the CNN
model and their standardised spectral signature}

\end{figure}%

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\subsection{Intertidal vegetation
mapping}\label{intertidal-vegetation-mapping}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\phantomsection\label{cell-fig-workflow}
\begin{figure}[H]

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{./Figures/Figure3_workflow.png}

}

\caption{\label{fig-workflow}Schematic representation of the workflow.
Diamonds represent input or output data, and rectangles represent Python
processing algorithms. The overall workflow of this study is divided
into two distinct parts based on the spatial resolution of the drone
flights: high-resolution flights were utilized for training and
prediction of the CNN model, whereas low-resolution flights were solely
employed for prediction purposes.}

\end{figure}%

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

The spectral similarities of reflectance signatures between intertidal
green macrophytes (Magnoliopsida and Chlorophyceae) make their
discrimination challenging using simple classification algorithms
(Figure~\ref{fig-vegetation} F). To overcome this challenge, a deep
learning classification method was developed, trained, validated, and
applied to each drone flight (Figure~\ref{fig-workflow}).

\subsubsection{Neural Network model
building}\label{neural-network-model-building}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\global\setlength{\Oldarrayrulewidth}{\arrayrulewidth}

\global\setlength{\Oldtabcolsep}{\tabcolsep}

\setlength{\tabcolsep}{0pt}

\renewcommand*{\arraystretch}{1.5}

\providecommand{\ascline}[3]{\noalign{\global\arrayrulewidth #1}\arrayrulecolor[HTML]{#2}\cline{#3}}

\begin{longtable}[c]{ccc}
\caption{Vegetation Classes of the model and the number of pixels used to train
and validate each class}\tabularnewline

\ascline{1.5pt}{666666}{1-3}

\multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Class}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Training\ Pixels}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Validation\ Pixels}}} \\

\ascline{1.5pt}{666666}{1-3}\endfirsthead

\ascline{1.5pt}{666666}{1-3}

\multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Class}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Training\ Pixels}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Validation\ Pixels}}} \\

\ascline{1.5pt}{666666}{1-3}\endhead

\multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Bacillariophyceae}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{4,475}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{3,371,920}}} \\

\multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Chlorophyceae}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{17,140}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{6,258,737}}} \\

\multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Magnoliopsida}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{221,065}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{69,079,189}}} \\

\multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Phaeophyceae}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{169,936}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{18,481,141}}} \\

\multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Rhodophyceae}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{5,771}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{}}} \\

\ascline{1.5pt}{666666}{1-3}

\end{longtable}

\arrayrulecolor[HTML]{000000}

\global\setlength{\arrayrulewidth}{\Oldarrayrulewidth}

\global\setlength{\tabcolsep}{\Oldtabcolsep}

\renewcommand*{\arraystretch}{1}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

A dataset containing photo-interpreted drone reflectance pixels was
built to train a Convolutional Neural Network (CNN) model with 2 hidden
layers. The training pixels were categorized into 7 different classes,
representing the various habitats encountered at the different study
sites: Sediment, Water, Chlorophyceae, Magnoliopsida, Bacillariophyceae,
Phaeophyceae and Rhodophyceae. Only low-altitude flights (Table 1) were
used for training purposes because of their 8 mm spatial resolution
allowing to avoid spectral sub-pixel mixing and to accurately visually
differentiate various vegetation classes. More than 418,000 pixels at 8
mm resolution from the 3 training flights have been used to train the
model (Table 2). Twenty one variables were used by the model as
predictors: 10 raw spectral bands of the Micasense RedEdge Dual MX
multispectral camera (ranging from 444 nm to 840 nm), the same 10
spectral bands standardized using a min/max transformation
(Equation~\ref{eq-std} ; \citep{cao2017}) and the NDVI
(Equation~\ref{eq-ndvi}). Standardisation of spectral bands is used to
eliminate the scaling differences between spectra and to limit the
effect of biomass on the shape of the spectra \citetext{\citealp[
]{Douay2022}; \citealp{Davies2023}}.

\begin{equation}\phantomsection\label{eq-std}{
R_{i}^{*}(\lambda) = \frac{R_{i}(\lambda) - min(R_{i})}{max(R_{i})- min(R_{i})}
}\end{equation}

where \(R_{i}(\lambda)\) is the reflectance at the wavelength
\((\lambda)\) of each individual spectra \((i)\), \(min(R_{i})\), and
\(max(R_{i})\) are the minimum and maximum value of the spectra \((i)\)

\begin{equation}\phantomsection\label{eq-ndvi}{
NDVI = \frac{R(840nm)-R(668nm)}{R(840nm)+R(668nm)}
}\end{equation}

where \(R(840nm)\) is the reflectance at 840 nm and \(R(668nm)\) is the
reflectance at 668 nm.

\subsubsection{Validation}\label{validation}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

The CNN model was applied to all the 9 flights at both 12 and 120 m of
altitude. In situ information on georeferenced class type and percent
cover collected during each flight has been used to assess the CNN
accuracy. These images were used to construct a validation dataset
indicating the presence or absence of each class. Additionally to the
quadrat-based validation dataset, polygons of each class were photo
interpreted in order to increase the number of pixels of the validation
dataset. A confusion matrix, along with precision metrics such as the
global accuracy, the sensitivity, the specificity, and the Kappa
coefficient, was generated for each of the study sites. Finally each
validation matrix has been merge to create a unique matrix of the model.
A total of 536,000 pixels has been used to globally validate the CNN
model.

\section{Results}\label{results}

\subsection{Classification}\label{classification}

A total of 9 prediction maps corresponding to the 9 drone flights were
obtained. Each prediction map is associated with a probability map,
indicating the probability of the selected class for every pixel. The
low-altitude flight conducted in Gafanaha, Portugal, represents the site
with the highest complexity (Figure~\ref{fig-GafLow}). Among the 5
vegetation classes on which the CNN model was trained, 4 were present on
this site. On this site, there is a mixture of Chlorophyceae and
Rhodophyceae over the seagrass meadow. This is also where
Bacillariophyceae is most abundant. Although the seagrass bed is solely
composed of \emph{Zostera noltei}, various colors can be observed: dark
green (indicating healthy beds with 100\% coverage) and whitish/brown
(indicating beds where leaves are dying). Regardless of the color, the
meadow is predicted as Magnoliopsida by the CNN model.

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\phantomsection\label{cell-fig-GafLow}
\begin{figure}[H]

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{./Figures/Maps Pred/FigX-Gaf_Low_Pred.png}

}

\caption{\label{fig-GafLow}RGB orthomosaic (Left) and Prediction (Right)
of the low altitude flight of Gafanaha, Portugal. The total extent of
this flight is 3000m² with a resolution of 8 mm per pixel. Background
colors means intertidal area (Light Green) and land area (Light Grey).
The zoom covers an area equivalent to a 10-meter Sentinel-2 pixel size.}

\end{figure}%

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

The high-altitude flight over Gafanha covered nearly 1 km² in total
(Figure~\ref{fig-GafHigh}). A channel delineating a small island was
present in the flight, but it has been masked in the prediction map due
to glinting and misclassification caused by the turbid shallow water.
Most of the intertidal area has been classified as Magnoliopsida by the
model, even though there is discoloration of seagrass blades. Only a few
pixels have been classified as Chlorophyceae at this scale. Furthermore,
the area that was classified as Bacillariophyceae in the low-altitude
flight remains mostly classified as such in the high-altitude flight,
but some pixels of it are now classified as Magnoliopsida. The patches
of Rhodophyceae are still well classified. \emph{Spartina sp.} was
present in the northern part of the flight, near the land, and has been
classified either as Magnoliopsida or as Phaeophyceae.

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\phantomsection\label{cell-fig-GafHigh}
\begin{figure}[H]

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{./Figures/Maps Pred/FigX-Gaf_High_Pred.png}

}

\caption{\label{fig-GafHigh}RGB orthomosaic (Left) and Prediction
(Right) of the high altitude flight of Gafanaha, Portugal. The total
extent of this flight is about 1 km² with a resolution of 80 mm per
pixel. Background colors means intertidal area (Light Green), land area
(Light Grey) and water (Light Blue). The red triangle shows the extent
of the low altitude flight of Gafanha. The zoom covers an area
equivalent to a 10-meter Sentinel-2 pixel size.}

\end{figure}%

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

The High altitude flight made in the inner lagoon of the Ria de Aveiro
is the largest of all flights, covering almost 1.5 km²
(Figure~\ref{fig-Boat}). On this site, only seagrass and red algae where
seen on the field. The classification follows the same pattern, with a
patchy Magnoliopsida meadow mixed on the eastern part of the flight with
Rhodophyceae. As shown in the zoom (Figure~\ref{fig-Boat}), the edges of
the meadow can be colonised by Chlorophyceae.

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\phantomsection\label{cell-fig-Boat}
\begin{figure}[H]

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{./Figures/Maps Pred/FigX-Boat_Pred.png}

}

\caption{\label{fig-Boat}RGB orthomosaic (Top) and Prediction (Bottom)
of the flight made in the inner part of Ria de Aveiro Lagoon, Portugal.
The total extent of this flight is about 1.5 km² with a resolution of 80
mm per pixel. Background colors means intertidal area (Light Green),
land area (Light Grey) and water (Light Blue). The zoom covers an area
equivalent to a 10-meter Sentinel-2 pixel size.}

\end{figure}%

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

The flight over L'Epine in Noirmoutier Island, France
(Figure~\ref{fig-Dike}) was conducted near a dike crossing the northern
part of the flight from west to east. Alongside this dike, brown algae
attached to rocks and stranded green algae could be found. Despite the
high mixture between Chlorophyceae and Magnoliopsida in this flight,
these two classes are well discriminated by the classifier.

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\phantomsection\label{cell-fig-Dike}
\begin{figure}[H]

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{./Figures/Maps Pred/FigX-Dike_Pred.png}

}

\caption{\label{fig-Dike}RGB orthomosaic (Top) and Prediction (Bottom)
of Northern part of Noirmoutier Island, France. The total extent of this
flight is about 28 000 m² with a resolution of 80 mm per pixel.
Background colors means intertidal area (Light Green) and land area
(Light Grey). The zoom covers an area equivalent to a 10-meter
Sentinel-2 pixel size.}

\end{figure}%

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\subsection{Validation}\label{validation-1}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\phantomsection\label{cell-fig-Validation}
\begin{figure}[H]

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{./Figures/Validation/ConfusionMatrixGlobal.png}

}
